model-understanding

### 2019-05-05 stats/model-understanding.Rmd
simple model for effective range of item parameters; some viz

```{r}
library(tidyverse)
library(rwebppl)
library(brms)
library(coda)
library(ggthemes)
library(readxl)
library(ggpubr)
library(stringr)

@@

```{r old}
semKnowPriors <- '

@@

  var globalLineParameters = {
    intercept:  discreteUniformBins(0, 5, 2),
    slope: discreteUniformBins(-2, 2, 2)
  }

  var itemVariability = {
    intercept: discreteUniformBins(0.01, 5, 2),
    slope: discreteUniformBins(0.01, 5, 2)
  }
  // conceptually it might not make sense to have two sigmas, but maybe it does
  // intercept sigma probably will be smaller than slope sigma

  // display([
  //  globalLineParameters.intercept, itemVariability.intercept,
  //  globalLineParameters.slope, itemVariability.slope,
  //  itemLineParameters.intercept, itemLineParameters.slope
  // ])
  var output = map(function(globalLineIntercept){
    map(function(globalLineSlope){
      map(function(itemVariabilityIntercept){
          map(function(itemVariabilitySlope){

            var itemLineParameters = {
              intercept: discreteUniformBins(
                globalLineIntercept - 1*itemVariabilityIntercept,
                globalLineIntercept + 1*itemVariabilityIntercept,
                itemVariabilityIntercept
              ),
              slope: discreteUniformBins(
                globalLineSlope - 1*itemVariabilitySlope,
                globalLineSlope + 1*itemVariabilitySlope,
                itemVariabilitySlope
              )
            }

            map(function(itemLineIntercept){
                map(function(itemLineSlope){
                  map(function(age){
                    var sem_knowledge = logistic(itemLineIntercept + 
                      itemLineSlope * age)
                    // display(itemLineParameters.intercept + "+ " + itemLineParameters.slope + " X " + age + " =logistic= " + sem_knowledge)
                    var priorProbs =  [.5,.5]
                    var speakerOptimality = 1

                    var modelPredictions = pragmaticListener({label: "novel_word"}, 
                            priorProbs, speakerOptimality, sem_knowledge)
                    var smk =  "semKnowledge_" + age
                    var mpa =  "modelPrediction_" + age
                    return {
                      global_intercept: globalLineIntercept,
                      global_slope: globalLineSlope,
                      sigma_intercept:itemVariabilityIntercept, 
                      sigma_slope:itemVariabilitySlope, 
                      item_intercept: itemLineIntercept,
                      item_slope: itemLineSlope,
                      age: age,
                      sem_knowledge : sem_knowledge,
                      L1 : Math.exp(modelPredictions.score(1))
                    }

@@

  _.flattenDeep(output)
  // return query
}
'
```

@@

# copied code

=>
```{r}
save(sem_know_model, "parameter_enumeration_so1_flatPriors.RData")
load(file = "parameter_enumeration_so1_flatPriors.RData")# sem_know_model
```

```{r}
baseMeModel <- '

var sem_knowledge = 1

var speakerOptimality = 1

modelPredictions

```{r}
webppl(program_code = paste(rsaUtils, rsaModel, baseMeModel,sep='\n'))
```

# Inferring semantic knowledge
```{r}
semStr <- '
var data = dataFromR.data

@@

sem_know_model<- webppl(
  program_code = paste(rsaUtils, rsaModel, semStr , sep='\n'),
  data = list(data = me_model_data),
  data_var = "dataFromR",
  model_var = "model",
  chains = 3,
  cores = 3,
  inference_opts = list(method = "incrementalMH", 
                        samples = 2000, 
                        burn = 1000,
                        verbose = T,
                        verboseLag = 500)
)

#saveRDS(sem_know_model, "../saves/sem_know_model.rds")
sem_know_model <- readRDS("../saves/sem_know_model.rds")
```

>>>>

```{r}
library(tidyverse)
library(viridis)
library(rwebppl)
library(brms)
library(coda)
library(ggthemes)
library(readxl)
library(ggpubr)
library(stringr)

@@

```{r loop through bda model structure}
semKnowPriors <- '

@@

  var globalLineParameters = {
    intercept:  discreteUniformBins(0, 5, 1),
    slope: discreteUniformBins(-2, 2.01, 1)
  }

  var itemVariability = {
    intercept: discreteUniformBins(0.01, 5, 1),
    slope: discreteUniformBins(0.01, 5, 1)
  }
  // conceptually it might not make sense to have two sigmas, but maybe it does
  // intercept sigma probably will be smaller than slope sigma

  // display([
  //  globalLineParameters.intercept, itemVariability.intercept,
  //  globalLineParameters.slope, itemVariability.slope,
  //  itemLineParameters.intercept, itemLineParameters.slope
  // ])
  var output = map(function(globalLineIntercept){
  display(globalLineIntercept)
    map(function(globalLineSlope){
      map(function(itemVariabilityIntercept){
          map(function(itemVariabilitySlope){

            var itemLineParameters = {
              intercept: discreteUniformBins(
                globalLineIntercept - 2*itemVariabilityIntercept,
                 globalLineIntercept + 2*itemVariabilityIntercept,
                 itemVariabilityIntercept
              ),
              slope: discreteUniformBins(
                 globalLineSlope - 2*itemVariabilitySlope,
                globalLineSlope + 2*itemVariabilitySlope,
                itemVariabilitySlope
              )
            }

            map(function(itemLineIntercept){
                map(function(itemLineSlope){
                  map(function(age){
                    var sem_knowledge = logistic(
                      itemLineIntercept + itemLineSlope * age
                    )

                    var priorProbs =  [.5,.5]
                    var speakerOptimality = 1

                    return {
                      global_intercept: globalLineIntercept,
                      global_slope: globalLineSlope,
                      sigma_intercept:itemVariabilityIntercept, 
                      sigma_slope:itemVariabilitySlope, 
                      item_intercept: itemLineIntercept,
                      item_slope: itemLineSlope,
                      age: age,
                      sem_knowledge : sem_knowledge,
                      L1 : Math.exp(modelPredictions.score(1))
                    }

@@

  _.flattenDeep(output)
}
'
```

@@

subset.ages <- unique(sem_know_model$age)[seq(1, 90, 10)]
```

Parameter ranges

```{r}
summary(sem_know_model)
```

- Is it expected that L1 predicitons should only range between 0.25 and 0.75? (this may be because speaker optimality = 1)

### L1 prediction vs. sem_knowledge

```{r}
sem_know_model %>% 
  distinct(sem_knowledge, L1) %>%
  ggplot(., aes( x = sem_knowledge, y = L1 ))+
  geom_point()+
  ylim(0, 1)
```

### L1 prediction vs. item intercept

```{r}
sem_know_model %>% 
  distinct(item_intercept, L1) %>%
  ggplot(., aes( x = item_intercept, y = L1 ))+
  geom_point()+
  ylim(0, 1)
```

### L1 prediction vs. item slope

```{r}
sem_know_model %>% 
  distinct(item_slope, L1) %>%
  ggplot(., aes( x = item_slope, y = L1 ))+
  geom_point()+
  ylim(0, 1)
```

#### Original priors entailed the effective range for item parameters of [-10, 10] for the slope and [-8, 8] for the intercept

```{r loop through relevant range}
semKnowPriorsReduced <- '

@@

var model  = function(){
  var itemLineParameters = {
    intercept: discreteUniformBins( -8, 9, 1),
    slope: discreteUniformBins( -10, 11, 1 )
  }
  var output = map(function(itemLineIntercept){
                map(function(itemLineSlope){
                  map(function(age){
                    var sem_knowledge = logistic(
                      itemLineIntercept + itemLineSlope * age
                    )

                    var priorProbs =  [.5,.5]
                    var speakerOptimality = 1

                    return {
                      item_intercept: itemLineIntercept,
                      item_slope: itemLineSlope,
                      age: age,
                      sem_knowledge : sem_knowledge,
                      L1 : Math.exp(modelPredictions.score(1))
                    }

  _.flattenDeep(output)
}
'
```

```{r}

sem_know_model_reduced <- webppl(
  program_code = paste(rsaUtils, rsaModel, semKnowPriorsReduced , "model()", sep='\n'),
  data = list(data = me_model_data),
  data_var = "dataFromR",
  model_var = "model",
  #inference_opts = list(method = "enumerate")
)

subset.ages <- unique(sem_know_model$age)[seq(1, 90, 10)]
```

### L1 prediction vs. item slope X item intercept

```{r}
sem_know_model_reduced %>% 
  filter(age %in% subset.ages) %>%
  filter(item_intercept < 5, item_intercept > -5,
         item_slope < 5, item_slope > -5) %>%
  distinct(age, item_intercept, item_slope, sem_knowledge) %>%
  ggplot(., aes( x = item_intercept, y = item_slope, fill = sem_knowledge ))+
  geom_tile()+
  facet_wrap(~age)+
  scale_fill_viridis()

ggsave("semKnowledge_bySlope_int.pdf")
```
